% This document uses the custom MyCV class
% This document must be compiled using XeLaTeX!

% Set documentclass
\documentclass[11 pt, a4paper]{MyCV}



% Define commands for personal data
\newcommand{\mobilePhone}{\textcolor{darkCerulean}\faMobile ~~ (+49) 0175-4134389 (mobile)}
\newcommand{\privatePhone}{\textcolor{darkCerulean}\faPhone ~~ (+49) 040-6306713 (private)}
\newcommand{\officePhone}{\textcolor{darkCerulean}\faPhoneSquare ~~ (+49) 040-8998-2249 (office)}

\newcommand{\privateMail}{\textcolor{darkCerulean}\faEnvelope ~~ benedict.kaiser@live.de (private)}
\newcommand{\officeMail}{\textcolor{darkCerulean}\faEnvelopeSquare ~~ benedict.kaiser@desy.de (office)}

\newcommand{\separator}{~~ $\mid$ ~~}


\begin{document}

%--------------------------
% HEADER
%--------------------------

\begin{center}
	\begingroup
		\HUGE
		\headerfontlight{Benedict} \headerfontbold{Kaiser}
	\endgroup
	\\
	\begin{footnotesize}
		\headerfontlight{\textit{Steilshooper Straße 328, 22309 Hamburg, Federal Republic of Germany}}
	\end{footnotesize}
	\\[0.5em]
	\begin{scriptsize}
		\mobilePhone \separator \privatePhone \separator \officePhone\\
		\privateMail \separator \officeMail
	\end{scriptsize}
\end{center}

\vspace*{2cm}


%--------------------------
% RECIPIENTS ADDRESS
%--------------------------

\begingroup
	\bodyfont
	Prof. Dr. Dieter Horns \hfill \textit{\themainfont\today} \\
	– Vorsitzender der Auswahlkommission Physik M.Sc. –
\endgroup

Universität Hamburg\\
Fachbereich Physik\\
Luruper Chaussee 149, Gebäude 68 , Raum 13\\
22761 Hamburg\\

\vspace*{0.75cm}

%--------------------------
% SUBJECT
%--------------------------

\begingroup
	\bodyfont
	\underline{Application for a master degree course in physics}
\endgroup


%--------------------------
% MAIN BODY
%--------------------------
Dear selection panel "Physics M. Sc."\\
Dear Prof. Dr. Horns,

it is with great pleasure and anticipation that I am submitting you my application for a Master degree course in physics at the University of Hamburg starting the upcoming semester (winter semester 2016/17).

I am currently enrolled for a bachelor degree course of physics in Hamburg, where I am now working on my Bachelor thesis in a research group that works in the area of experimental particle physics and the upcoming Jiangmen Underground Neutrino Observatory (JUNO). There, my research work focusses mainly on programming and implementing control software and monitoring systems for photomultipliers that will be needed in the JUNO experiment.

My studies there have already made me familiar with working in laboratories and gave me a good insight into working in a scientific research group. In particular, they reinforced my wish to further specialize in this topic, as I find it especially interesting to contribute to a vast physics project in an international collaboration. I strongly believe that my strong research and language skills suit the need of a member of a research group in physics.

It might be appropriate to mention, that I am planning on continuing to work as a student assistant in the upcoming semester. I have been fortunate enough to already recieve an offer in another particle physics group. 

I would be honored to be given the opportunity to further deepen my knowledge in physics at the University of Hamburg, as specialized institutions such as the Center for Free-Electron Laser Science (CFEL) and the close collaboration with DESY provide a vast field of opportunities to work internationally on a research project. I do however, also have personal and familiar reasons that make me wish to continue living and studying in Hamburg. I am certain, that the University of Hamburg will be the right place to continue my studies.

%I am choosing to deepen my knowledge in physics at the University of Hamburg, because specialized institutions such as the Center for Free-Electron Laser Science (CFEL) and the close collaboration with DESY provide a vast field of opportunities to work internationally on a research project. I do however, also have personal and familiar reasons that make me wish to continue studying in Hamburg. I am certain, that the University of Hamburg will be the right place to continue my studies.

I enclose a European curriculum vitae, a language certificate as well as a certificate for an international competence course and the required forms for a application at the University of Hamburg.

Please do not hesitate to contact me, should you stil have further questions or need more documents. I highly appreciate you taking your time looking through my application and I look forward to hearing from you.

Yours faithfully,

\vspace*{1.5cm}

Benedict Kaiser

\end{document}