%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE
%----------------------------------------------------------------------------------------
\lettersection{\faBriefcase \hspace{0.1cm} Relevant Professional Experience}

%----------------------------------------------------------------------------------------
%	IPsoft Technical Executive
%----------------------------------------------------------------------------------------
\begin{tabularx}{1\linewidth}{>{\raggedleft\scshape}p{2.5cm}X}
\gray Employer & \textbf{IPsoft} \hfill Amsterdam, The Netherlands\\
\gray Job Title & \textbf{Technical Executive} \hfill May 2015 --- now\\
\gray Reference & \textbf{Nick Walsh} \\
\end{tabularx}

\vspace{2pt}
In the position of Technical Executive I am responsible for overseeing the day to day IT operations for some of our top clients. To ensure that the organization’s IT infrastructure works efficiently I strongly focused on identifying waste and eliminating it from the IT processes. By automating standard changes, deployments and a large part of the incident management process the operational FTE requirements are strongly reduced while client satisfaction is increased. The global, multi-site IT environment consists of over 1000 Windows and Linux/Unix servers running on top of a combination of Azure solutions and on-premise vSphere environments.\\
\\
\textbf{Key Contributions:}
\begin{itemize}
\item Automated Cloud deployments using Azure Resource Templates and Azure Automation DSC.
\item Created a build, test and deploy pipeline for Windows Monitoring checks using PowerShell, Pester and Bamboo.
\item Leading a team of 7 Senior IT engineers by training, motivating and setting clear objectives.
\item Building a weekly automation pipeline assuring maximum autonomic resolution for the client.
\item Build and bug-fix over a hundred automatons which either diagnose or completely resolve recurring issues saving my employer multiple FTE's. Used technologies include JavaScript, PowerShell, Bash and MySQL. 
\item Worked with tier 1 clients to align and support their requirements with a high level of attention and intimacy.
\item Leading technical projects, system and software upgrades to improve the client's on-premise infrastructure. 
\item Building a strong trust and positive relationship with the client.
\item Created over fifty automated SOP’s to ensure uniform effective processing of standard changes and incidents.
\item Manage relationships with third party vendors for technical products and services for the organization.\\
\end{itemize}

%----------------------------------------------------------------------------------------
%	IPsoft Automation Engineer.
%----------------------------------------------------------------------------------------
\begin{tabularx}{1\linewidth}{>{\raggedleft\scshape}p{2.5cm}X}
\gray Employer & \textbf{IPsoft} \hfill Amsterdam, The Netherlands\\
\gray Job Title & \textbf{Automation Engineer} \hfill February 2017 --- now (Dual role)\\
\gray Reference & \textbf{Michell Grauwmans} \\
\end{tabularx}
By automating repetitive IT-processes IPsoft's Automation as a Service (AaaS) enables customers to assume a more proactive role towards improving their IT infrastructure. In my role as Automation Engineer I'm responsible for maintaining a close relationship with customers, partners and engineers. By translating standard operating procedures to automated workflows we streamline the operational processes of our customers.\\

\textbf{Key Contributions:}
\begin{itemize}
\item Setup Jira, Confluence and Bitbucket to manage the automation pipeline, documentation and code. 
\item Developed a tool in JavaScript to automate the initialization and validation of connections between our IPcenter platform and our client's Windows/Linux infrastructure. 
\item Automated incident management pipelines by integrating cmdb, incident- and password management tooling with our automation platform by utilizing different API's. 
\item Setup a Instructor-led Training Program to on-board new staff.
\end{itemize}

%----------------------------------------------------------------------------------------
%	Other experience
%----------------------------------------------------------------------------------------
\lettersection{\faBackward \hspace{0.1cm} Previous Professional Experience (2005 - 2015)}
My full IT career dates back to 2005, hence I made an effort to keep my resume short and relevant. A full detailed history of my professional experience is available on my {\faLinkedIn} \url{linkedin.com/in/JavydeKoning} profile.