\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{References}

\bibitem[Aravind, Hackl \& Wexler(2017)]{aravind2017}
Aravind, A.\,, Hackl, M.\, \& Wexler, K.\,, 2017, Syntactic and Pragmatic Factors in Children's Comprehension of Cleft Constructions [J], \emph{Langauge Acquisition}, 1: 1-31.
\bibitem[Beaver \& Clark(2008)]{beaver2008}
	Beaver, D.\,I.\,\& Clark, B.\,Z.\,, 2008, \emph{Sense and Sensitivity: How Focus Determines Meaning} [M], Malden, MA: Blackwell Publishing.
\bibitem[Brandt et al(2014)]{brandt2014}
	Brandt, M.\,et al, 2014, The Replication Recipe: What makes for a convincing replication? [J], \emph{Journal of Experimental Social Psychology}, 50: 217--224.
\bibitem[Colonna, Schimke \& Hemforth(2012)]{colonna2012}
	Colonna, S.\,, Schimke, S.\,, \& Hemforth, B.\,, 2012, Information structure effects on anaphora resolution in German and French: A crosslinguistic study of pronoun resolution [J], \emph{Linguistics}, 55(5): 991--1013.
\bibitem[Colonna, Schimke \& Hemforth(2015)]{colonna2015}
	Colonna, S.\,, Schimke, S.\,, \& Hemforth, B.\,, 2015, Deifferent effects of focus in intra- and inter-sentential pronoun resolution in German [J], \emph{Language, Cognition and Neuroscience}, 30(10): 1306--1325.
\bibitem[Cowles, Walenski \& Kluender(2007)]{cowles2007}
	Cowles, H.\,H.\,, Walenski, M.\,, \& Kluender, R.\,, 2007, Linguistic and cognitive prominence in anaphora resolution: topic, contrastive focus and pronouns [J], \emph{Topoi}, 26(1): 3--18.
\bibitem[Grosz, Weinstein \& Joshi(1995)]{grosz1995}
	Grosz, B.\,J.\,, Weinstein, S.\,, \& Joshi, A.\,K.\,, 1995, Centering: A framework for modeling the local coherence of discourse [J], \emph{Computational Linguistics}, 21: 203--205.
\bibitem[Engelkamp \& Zimmer(1982)]{engelkamp1982}
	Engelkamp, J.\, \& Zimmer, H.\,D.\,, 1982, The interaction of subjectivization and concept placement in the processing of cleft sentences [J], \emph{Quarterly Journal of Experimental Psychology: Human Experimental Psychology}, 34(A): 463--478.
\bibitem[Hakes, Evans \& Brannon(1976)]{hakes1976}
	Hakes, D.\,T.\,, Evans, J.\,S.\, \& Brannon, L.\,L.\,, 1976, Understanding sentences with relative clauses [J], \emph{Memory and Cognition}, 4(3): 283--290.
\bibitem[Halliday(1967)]{halliday1967}
	Halliday, M.\,, 1967, Notes on Transitivity and Theme in English (Part 2) [J], \emph{Journal of Lingyuistics}, 3: 206.
\bibitem[Kaiser(2011)]{kaiser2011}
	Kaiser, E.\,, 2011, Focusing on pronouns: Consequences of subjecthooh, pronominalisation and contrastive focus [J], \emph{Language and cognitive Processes}, 26(10), 1625--1666.
\bibitem[Kornai(2008)]{kornai2008}
	Kornai, A.\,, 2008, \emph{Mathematical Linguistics} [M], London: Springer-Verlag. 
\bibitem[Lambrecht(2001)]{lambrecht2001}
	Lambrecht, K.\,, 2001, A framework for the analysis of cleft constructions [J], \emph{Linguistics}, 39(3): 463--516.
\bibitem[Liu, Xu \& Liang(2017)]{liu2017}
	Liu, H.\,, Xu, C.\, \& Liang, J.\,, 2017, Dependency distance: A new perspective on syntactic patterns in natural langauges [J], \emph{Physics of Life Reviews}, 21: 171--193.
\bibitem[Manning \& Sch\"{u}tze(1999)]{manning1999}
	Manning, C.\,D.\, \& Sch\"{u}tze, H.\,, 1999, \emph{Foundations of Statistical Natural Language Processing} [M], London: The MIT Press. 
\bibitem[Mitkov(1999)]{mitkov1999}
	Mitkov, R.\,, 1999, \emph{Anaphora Resolution: The State of the Art} [R], Paper based on the COLING'98/ACL'98 tutorial on anaphora resolution.
\bibitem[O'Grady et al(2010)]{ogrady2010}
	O'Grady, W.\, et al, 2010, \emph{Contemporary Linguistics: An Introduction} [M], Boston: Bedford \& St. Martin's.
\bibitem[Patterson, Esaulova \& Felser(2017)]{patterson2017}
	Patterson, C.\,, Esaulova, Y.\, \& Felser, C.\,, 2017, The impact of focus on pronoun resolution in native and non-native sentence comprehension [J], \emph{Second Langauge Research}, 33(4): 403--429.
\bibitem[Pinker(2014)]{pinker2014}
	Pinker, S.\,, 2014, \emph{The Sense of Style: The Thinking Person's Guide to Writing in the 21$^{st}$ Century} [M], New York: Penguin Books.
\bibitem[Reichle(2014)]{reichle2014}
	Reichle, R.\,, 2014, Cleft type and focus structure processing in French, \emph{Language, Cognition and Neuroscience}, 29(1): 107--124.
\bibitem[Sayed(2003)]{sayed2003}
	Sayed, I.\,Q.\,, 2003, \emph{Issues in Anaphora Resolution} [OL], retrieved from: \url{https://nlp.stanford.edu/courses/cs224n/2003/fp/iqsayed/project_report.pdf}.
\bibitem[Sgall, Hajicová \& Panevová(1986)]{sgall1986}
	Sgall, P.\,, Hajicová, E.\, \& Panevová, J.\,, 1986, \emph{The Meaning of the Sentence in its Semantic and Pragmatic Aspects} [M], Praha: Academia.
\bibitem[徐晓东, 陈庆荣(2014)]{徐晓东2014}
	徐晓东, 陈庆荣, 2014, 汉语焦点信息影响代词回指的电生理机制 [J], 《心理科学进展》, 22(6): 902--910.
\bibitem[徐晓东，倪传斌，陈丽娟(2013)]{徐晓东2013}
	徐晓东，倪传斌，陈丽娟, 2013, 话题结构和动词语义对代词回指的影响：一项基于语言产生和语言理解任务的实证研究 [J], 《现代外语（季刊）》, 36(4): 331--339. 
\end{thebibliography}
