\documentclass[10pt]{article}
\usepackage[english]{babel}

\input{config/minimal-resume-config}


\begin{document}

% Personal (includes three rows: Name, Address, Contact)
% -----------------------------------------------------

\personal{
  % Name
  RATUL\hspace{\sizefour}SAHA
  }{
  % Address
  {13} COMPUTING DRIVE, SINGAPORE {117417}
  }{
  % contact
  www.ratulsaha.com\hspace{\sizethree}ratul@comp.nus.edu.sg
}

% Chapter: Education
% ------------------

\chap{EDUCATION}{

  \subchap{National University of Singapore, Singapore}{2013--present}{
    \begin{newitemize}
      \item PhD student in Department of Computer Science, School of Computing

      \item Thesis: \textit{Towards Quantitative Model Checking of Distributed Probabilistic Systems}

      \item Advisor: David S. Rosenblum, and previously, P. S. Thiagarajan (retired)
    \end{newitemize}
  }

  \subchap{Chennai Mathematical Institute, India}{2009--2012}{
    \begin{newitemize}
      \item B.Sc in Mathematics and Computer Science

      \item GPA: {7.45}/10 and CGPA for CS: {9.25}/10
    \end{newitemize}
  }
}

% Chapter: Publications
% ---------------------

\chap{PUBLICATIONS}{

  \begin{newitemize}
    \item {Ratul Saha}, Madhavan Mukund, and R. P. Jagadeesh Chandra Bose. \textit{Time-bounded Statistical Analysis of Resource-constrained Business Processes with Real-time Distributed Probabilistic Systems}. In Proc. of SETTA 2016.

    \item Javier Esparza, Philipp Hoffmann, {Ratul Saha}. \textit{Polynomial Analysis Algorithms for Free Choice Probabilistic Workflow Nets}. In Proc. of QEST 2016.

    \item Loi Luu, {Ratul Saha}, I. Parameshwaran, Prateek Saxena, and A. Hobor. \textit{On Power Splitting Games in Distributed Computation: The Case of Bitcoin Pooled Mining}. In Proc. of CSF 2015.

    \item {Ratul Saha}, Javier Esparza, Sumit K. Jha, Madhavan Mukund, and P. S. Thiagarajan. \textit{Distributed Markov Chains}. In Proc. of VMCAI 2015.
  \end{newitemize}
}

% Chapter: RESEARCH COLLABORATIONS
% ------------------------
\chap{RESEARCH COLLABORATIONS}{

\begin{newitemize}
  \item Collaborated with 10 professors, 3 PhD students, and 2 industry researchers across 8 cities

  \item Was invited as a guest researcher and speaker at

  \begin{newitemize}
    \item CFDVS lab, \textit{Indian Institute of Technology (IIT) Bombay, India} for Mar--April 2016. Host: S Akshay.

    \item ACMU, \textit{Indian Statistical Institute (ISI) Kolkata, India} for Jan--Feb 2016. Host: Ansuman Banerjee.

    \item CS Dept., \textit{Chennai Mathematical Institute (CMI), India} for Aug--Nov 2015. Host: Madhavan Mukund.

    \item FSR\&TCS, \textit{Technische Universit\"{a}t M\"{u}nchen (TUM), Germany} for May--July 2015. Host: Javier Esparza.
  \end{newitemize}

  \item Was invited to \textit{Trends and Challenges in Quantitative Verification}, A Mysore Park Workshop (2016)
\end{newitemize}
}

% Chapter: Awards
% ---------------

\chap{AWARDS}{

  \begin{newitemize}
    \item Among top 35 in the country to clear \textit{Indian National Mathematical Olympiad 2007}

    \item Was selected twice (2007 and 2008) to attend the training for selection of the Indian team for the International Mathematical Olympiad (IMO)

    \item Recipient of \textit{Undergraduate Scholarship} as part of the Nurture Program by National Board for Higher Mathematics, India
  \end{newitemize}
}

% NOTE: Forcing the content in a new page is optional
\newpage

% Chapter: Work Experience
% ------------------------

\chap{WORK EXPERIENCE}{

  \subchap{National University of Singapore, Singapore}{Jan--July 2013}{

    \begin{newitemize}
      \item Research assistant in School of Computing

      \item Worked on a model checking tool (using Python) for models with uncertainty
    \end{newitemize}
  }

  \subchap{LimberLink Technologies Pvt. Ltd., India}{Jun--July 2012}{

  \begin{newitemize}
    \item Software development internship under supervision of V. Vinay

    \item Designed a product to help educators provide affordable online education
  \end{newitemize}
  }

  \subchap{Chennai Mathematical Institute, India}{2010, 2012}{

    \begin{newitemize}
      \item Placement co-ordinator for Fall '12
      \item Teaching Assistant for Advanced Programming (Fall '10) and Data Mining (Fall '12)
    \end{newitemize}
  }
}

% Chapter: Leadership Experience
% ------------------------------

\chap{LEADERSHIP EXPERIENCE}{

  \subchap{Branding Consultancy}{2016--present}{
    \begin{newitemize}
      \item Developing online brands for a total of 5 small businesses, entrepreneurs, and celebrities

      \item Delegating designers and content managers to provide hassle-free support to clients
    \end{newitemize}
  }

  \subchap{Overleaf, Writelatex limited}{2014--present}{
    \begin{newitemize}
      \item Acting as an external advisor, ``Advisor of the month'' for Oct 2015

      \item Organized seminars on introduction to document preparation for beginners
    \end{newitemize}
  }

  \subchap{Movie Database and Promotion Agency}{2010--2013}{
    \begin{newitemize}
      \item Created a partnership firm and profiled more than 2000 movies and 4000 artists

      \item Led 8 part-time members including content managers, journalists, and national award-winning authors

      \item Was offered one acqui-hire and one acquisition
    \end{newitemize}
  }
}

% Chapter: Technical Experience
% -----------------------------

\chap{TECHNICAL EXPERIENCE}{

  \begin{newitemize}
    \item Programming languages: Python, Haskell, C and Java

    \item Tools and Scripts: \LaTeX, HTML5/CSS3/JS, MS Excel

    \item Designed a popular open-source presentation template called \textit{Presento} that is clean, minimal and extensible
  \end{newitemize}
}

% Chapter: Other Information
% --------------------------

\chap{OTHER INFORMATION}{
  \begin{newitemize}
     \item Nationality: Indian, Age: 24 years

     \item Languages: English (full professional proficiency), Bengali and Hindi (native or bilingual proficiency)

     \item TOEFL iBT (2012): 103/120

     \item GRE (2012): Verbal -- 147, Quantitative -- 164, Analytical writing -- 3.5

     \item Hobbies: swimming, chess, User Experience (UX) design
  \end{newitemize}
}

\end{document}
